KidsAIR initial SUMs data work

### 2020-01-21 dustrak_data_cleaning.Rmd
data cleaning updates

# Save pm dataset as RDS and CSV
# write_rds(kids_pm, "Output/kids_pm_full_raw.rds")
# write_csv(kids_pm, "Output/kids_pm_full.csv")
```

@@

##### Add new variables to Kids PM data #####
kids_pm_new <- kids_pm_joined %>%  
  #filter(area == "AK") %>% 
  # apply correction factor to pm and change to ug
  mutate(pm = pm/1.65,
         pm = pm*1000) %>% 
  mutate(home_winter_id = as.character(home_winter_id),
         area = as.factor(area),
         pm_date = date,
         pm_time = time,
         pm = as.numeric(pm),
         day_of_week = weekdays(date)) %>% 
  unite("pm_datetime", date:time, sep = " ") %>% 
  #select(home_winter_id, area, pm_date, pm_time, pm_datetime, day_of_week, pm) %>% 
  arrange(area, home_winter_id, pm_date) %>% 
  group_by(area, home_winter_id) %>% 
  # new variables summarizing pm2.5 within each home_winter_id
  mutate(pm_total_observations = n(),
         # new datetime values to calculate sampling interval
         min_datetime = first(pm_datetime),
         max_datetime = last(pm_datetime),
         min_datetime = ymd_hms(min_datetime),
         max_datetime = ymd_hms(max_datetime),
         diff_datetime = interval(min_datetime, max_datetime),
         # total sampling interval, based on first and last datetimes
         pm_sample_interval = diff_datetime/86400,
         # expected number of observations with the sampling interval
         expected_obs = pm_sample_interval*24*60,
         # difference between actual and expected observations
         actual_vs_expected_diff = pm_total_observations - expected_obs,
         # indicator variable if actual vs expected obs > 10 or greater
         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
                                      actual_vs_expected_diff > -10, 1, 0),
         pm_mean = mean(pm),
         # breaking sampling period into days
         obs = row_number(),
         sampling_day = if_else(obs < 1441, 1, 0),
         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
  # calculate pm means for each sampling day
  group_by(area, home_winter_id, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm),
         pm_sampling_day = sampling_day) %>% 
  separate(starttime, c("trash1", "starttime"), sep = " ") %>% 
  separate(stoptime, c("trash2", "stoptime"), sep = " ") %>% 
  ungroup() %>% 
  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day, 
         -datetime, -trash1, -trash2)

# Save pm dataset as RDS and CSV
# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
# write_csv(kids_pm_new, "Output/kids_pm_new.csv")
```

@@

# Fix dusttrak sample start times
## Notes from above were used to make the following changes in the data
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new.rds")

@@

kids_pm_new <- kids_pm %>% 

@@

# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
```

# Further cleaning:
## Remove specific homes from PM dataset
## Remove extremely high values
## Change negative values
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new.rds")

kids_pm_new <- kids_pm %>%
  # filtering out specific homes where PM data is not suitable for analysis
  # removes 119,520 obs
         # YKC06_105: only 12 hours of data
  mutate(do_not_use = if_else(area == "AK" & home_winter_id == "39", 1, 0),
         # YKC06_313: sample hourly instead of every minute
         do_not_use = if_else(area == "AK" & home_winter_id == "89", 1, do_not_use),
         # CH_302: sample hourly instead of every minute
         do_not_use = if_else(area == "NN" & home_winter_id == "127", 1, do_not_use),
         # CH_304: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "1330", 1, do_not_use),
         # CH_306: sample hourly instead of every minute
         do_not_use = if_else(area == "NN" & home_winter_id == "134", 1, do_not_use),
         # CH_307: only 8 datapoints
         do_not_use = if_else(area == "NN" & home_winter_id == "135", 1, do_not_use),
         # TC_357: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "1510", 1, do_not_use),
         # CH_322: sample hourly instead of every minute
         do_not_use = if_else(area == "NN" & home_winter_id == "161", 1, do_not_use),
         # CH_325: 6300 obs are negative, down to -12.12. Median is negative.
         do_not_use = if_else(area == "NN" & home_winter_id == "165", 1, do_not_use),
         # CH_400: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "2150", 1, do_not_use),
         # CH_403: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "2210", 1, do_not_use),
         # CH_108: Only 300 datapoints and no other files found
         do_not_use = if_else(area == "NN" & home_winter_id == "29", 1, do_not_use),
         # CH_207: Only 400 datapoints; no other files found
         do_not_use = if_else(area == "NN" & home_winter_id == "58", 1, do_not_use),
         # TC_230: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "620", 1, do_not_use),
         # CH_223: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "840", 1, do_not_use),
         # TC_242: 1 day worth of data with lots of gaps
         do_not_use = if_else(area == "NN" & home_winter_id == "86", 1, do_not_use),
         # WMA277: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "1910", 1, do_not_use),
         # WMA305: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "2500", 1, do_not_use),
         # WMA306: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "2510", 1, do_not_use),
         # WMA358: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "2860", 1, do_not_use),
         # WMA448: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "3750", 1, do_not_use),
         # WMA200: 4700 negative values, down to -4.85. Median is negative
         do_not_use = if_else(area == "WMT" & home_winter_id == "152", 1, do_not_use),
         # WMA201: 5000 negative values, down to -10. Median negative. 17 hour gap 
         do_not_use = if_else(area == "WMT" & home_winter_id == "153", 1, do_not_use),
         # WMA603: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "5100", 
                              1, do_not_use)) %>% 
  filter(do_not_use != 1) %>% 
  # replace PM > 10,000 with NA (33 obs)
  mutate(pm = if_else(pm > 10000, 999999, pm)) %>% 
  replace_with_na(replace = list(pm = 999999)) %>% 
  # replace PM < 1 (lower DustTrak limit) with 0.5 (lower limit / 2) 
  # 291,490 obs less than 1; 56,684 obs less than 0 
  mutate(pm = if_else(pm < 1, 0.5, pm))

# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
```

# New variables to use for analysis
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new.rds")
cleaning_summary <- read_xlsx("Input/cleaning_summary.xlsx") %>% 
  select(area, home_winter_id, final_cleaning_pm, sensitivity_analysis_pm)

kids_pm_new <- kids_pm %>% 
  mutate(home_winter_id = as.numeric(home_winter_id)) %>% 
  left_join(cleaning_summary, by = c("area", "home_winter_id")) %>% 
  select(-samplingdate, -samplingtime, -pickupdate, -pickuptime,
         -dtmin, -dtmax, -dtaverage, -qaqccomments, -data_comments,
         -pm_sampling_day, -pm_mean_daily) %>% 
  # create sampling days based on datetime differences
  group_by(area, home_winter_id) %>% 
  mutate(pm_datetime_new = if_else(is.na(pm_datetime_new), 
                                   pm_datetime, pm_datetime_new),
         first_datetime_new = first(pm_datetime_new),
         first_datetime_new = ymd_hms(first_datetime_new)) %>% 
  mutate(datetime_diff = pm_datetime_new - first_datetime_new,
         sampling_day = if_else(datetime_diff <= 86400, 1, 0),
         sampling_day = if_else(datetime_diff > 86400 & datetime_diff <= 172800, 
                                2, sampling_day),
         sampling_day = if_else(datetime_diff > 172800 & datetime_diff <= 259200, 
                                3, sampling_day),
         sampling_day = if_else(datetime_diff > 259200 & datetime_diff <= 345600, 
                                4, sampling_day),
         sampling_day = if_else(datetime_diff > 345600 & datetime_diff <= 432000, 
                                5, sampling_day),
         sampling_day = if_else(datetime_diff > 432000 & datetime_diff <= 518400, 
                                6, sampling_day),
         sampling_day = if_else(datetime_diff > 518400 & datetime_diff <= 604800, 
                                7, sampling_day),
         sampling_day = if_else(datetime_diff > 604800 & datetime_diff <= 691200, 
                                8, sampling_day),
         sampling_day = if_else(datetime_diff > 691200 & datetime_diff <= 777600, 
                                9, sampling_day),
         sampling_day = if_else(datetime_diff > 777600 & datetime_diff <= 864000, 
                                10, sampling_day),
         sampling_day = if_else(datetime_diff > 864000, 
                                11, sampling_day)) %>% 
  # new vars for daily means and % expected observations for each day
  group_by(area, home_winter_id, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm),
         percent_daily_obs_pm = n()/1440) %>% 
  ungroup() %>% 
  arrange(area, home_winter_id, pm_datetime_new) %>% 
  select(-first_datetime, -time_diff, -do_not_use)

@@

# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
# write_csv(kids_pm_new, "Output/kids_pm_new.csv")
```

>>>>

# Save pm dataset as RDS 
# write_rds(kids_pm, "Output/kids_pm_full_raw.rds")
```

@@

##### Add new variables to Kids PM data #####
kids_pm_new1 <- kids_pm_joined %>%  
  #filter(area == "AK") %>% 
  # apply correction factor to pm and change to ug
  mutate(pm = pm/1.65,
         pm = pm*1000) %>% 
  mutate(home_winter_id = as.character(home_winter_id),
         area = as.factor(area),
         pm_date = date,
         pm_time = time,
         pm = as.numeric(pm),
         day_of_week = weekdays(date)) %>% 
  unite("pm_datetime", date:time, sep = " ") %>% 
  #select(home_winter_id, area, pm_date, pm_time, pm_datetime, day_of_week, pm) %>% 
  arrange(area, home_winter_id, pm_date) %>% 
  group_by(area, home_winter_id) %>% 
  # new variables summarizing pm2.5 within each home_winter_id
  mutate(pm_total_observations = n(),
         # new datetime values to calculate sampling interval
         min_datetime = first(pm_datetime),
         max_datetime = last(pm_datetime),
         min_datetime = ymd_hms(min_datetime),
         max_datetime = ymd_hms(max_datetime),
         diff_datetime = interval(min_datetime, max_datetime),
         # total sampling interval, based on first and last datetimes
         pm_sample_interval = diff_datetime/86400,
         # expected number of observations with the sampling interval
         expected_obs = pm_sample_interval*24*60,
         # difference between actual and expected observations
         actual_vs_expected_diff = pm_total_observations - expected_obs,
         # indicator variable if actual vs expected obs > 10 or greater
         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
                                      actual_vs_expected_diff > -10, 1, 0),
         pm_mean = mean(pm),
         # breaking sampling period into days
         obs = row_number(),
         sampling_day = if_else(obs < 1441, 1, 0),
         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
  # calculate pm means for each sampling day
  group_by(area, home_winter_id, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm),
         pm_sampling_day = sampling_day) %>% 
  separate(starttime, c("trash1", "starttime"), sep = " ") %>% 
  separate(stoptime, c("trash2", "stoptime"), sep = " ") %>% 
  ungroup() %>% 
  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day, 
         -datetime, -trash1, -trash2)

# Save pm dataset as RDS 
# write_rds(kids_pm_new1, "Output/kids_pm_new1.rds")
```

@@

# Fix dusttrak sample start times
## Notes from above were used to make the following changes in the data
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new1.rds")

@@

kids_pm_new2 <- kids_pm %>% 

@@

# write_rds(kids_pm_new2, "Output/kids_pm_new2.rds")
```

# Further cleaning:
## Remove specific homes from PM dataset
## Remove extremely high values
## Change negative values
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new2.rds")

kids_pm_new3 <- kids_pm %>%
  # filtering out specific homes where PM data is not suitable for analysis
  # removes 119,520 obs
         # YKC06_105: only 12 hours of data
  mutate(do_not_use = if_else(area == "AK" & home_winter_id == "39", 1, 0),
         # YKC06_313: sample hourly instead of every minute
         do_not_use = if_else(area == "AK" & home_winter_id == "89", 1, do_not_use),
         # CH_302: sample hourly instead of every minute
         do_not_use = if_else(area == "NN" & home_winter_id == "127", 1, do_not_use),
         # CH_304: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "1330", 1, do_not_use),
         # CH_306: sample hourly instead of every minute
         do_not_use = if_else(area == "NN" & home_winter_id == "134", 1, do_not_use),
         # CH_307: only 8 datapoints
         do_not_use = if_else(area == "NN" & home_winter_id == "135", 1, do_not_use),
         # TC_357: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "1510", 1, do_not_use),
         # CH_322: sample hourly instead of every minute
         do_not_use = if_else(area == "NN" & home_winter_id == "161", 1, do_not_use),
         # CH_325: 6300 obs are negative, down to -12.12. Median is negative.
         do_not_use = if_else(area == "NN" & home_winter_id == "165", 1, do_not_use),
         # CH_400: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "2150", 1, do_not_use),
         # CH_403: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "2210", 1, do_not_use),
         # CH_108: Only 300 datapoints and no other files found
         do_not_use = if_else(area == "NN" & home_winter_id == "29", 1, do_not_use),
         # CH_207: Only 400 datapoints; no other files found
         do_not_use = if_else(area == "NN" & home_winter_id == "58", 1, do_not_use),
         # TC_230: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "620", 1, do_not_use),
         # CH_223: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "840", 1, do_not_use),
         # TC_242: 1 day worth of data with lots of gaps
         do_not_use = if_else(area == "NN" & home_winter_id == "86", 1, do_not_use),
         # WMA277: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "1910", 1, do_not_use),
         # WMA305: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "2500", 1, do_not_use),
         # WMA306: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "2510", 1, do_not_use),
         # WMA358: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "2860", 1, do_not_use),
         # WMA448: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "3750", 1, do_not_use),
         # WMA200: 4700 negative values, down to -4.85. Median is negative
         do_not_use = if_else(area == "WMT" & home_winter_id == "152", 1, do_not_use),
         # WMA201: 5000 negative values, down to -10. Median negative. 17 hour gap 
         do_not_use = if_else(area == "WMT" & home_winter_id == "153", 1, do_not_use),
         # WMA603: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "5100", 
                              1, do_not_use)) %>% 
  filter(do_not_use != 1) %>% 
  # replace PM > 10,000 with NA (33 obs)
  mutate(pm = if_else(pm > 10000, 999999, pm)) %>% 
  replace_with_na(replace = list(pm = 999999)) %>% 
  # replace PM < 1 (lower DustTrak limit) with 0.5 (lower limit / 2) 
  # 291,490 obs less than 1; 56,684 obs less than 0 
  mutate(pm = if_else(pm < 1, 0.5, pm))

# write_rds(kids_pm_new3, "Output/kids_pm_new3.rds")
```

# New variables to use for analysis
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new3.rds")
cleaning_summary <- read_xlsx("Input/cleaning_summary.xlsx") %>% 
  select(area, home_winter_id, final_cleaning_pm, sensitivity_analysis_pm)

kids_pm_new4 <- kids_pm %>% 
  mutate(home_winter_id = as.numeric(home_winter_id)) %>% 
  left_join(cleaning_summary, by = c("area", "home_winter_id")) %>% 
  select(-samplingdate, -samplingtime, -pickupdate, -pickuptime,
         -dtmin, -dtmax, -dtaverage, -qaqccomments, -data_comments,
         -pm_sampling_day, -pm_mean_daily) %>% 
  # create sampling days based on datetime differences
  group_by(area, home_winter_id) %>% 
  mutate(pm_datetime_new = if_else(is.na(pm_datetime_new), 
                                   pm_datetime, pm_datetime_new),
         first_datetime_new = first(pm_datetime_new),
         first_datetime_new = ymd_hms(first_datetime_new)) %>% 
  mutate(datetime_diff = pm_datetime_new - first_datetime_new,
         sampling_day = if_else(datetime_diff <= 86400, 1, 0),
         sampling_day = if_else(datetime_diff > 86400 & datetime_diff <= 172800, 
                                2, sampling_day),
         sampling_day = if_else(datetime_diff > 172800 & datetime_diff <= 259200, 
                                3, sampling_day),
         sampling_day = if_else(datetime_diff > 259200 & datetime_diff <= 345600, 
                                4, sampling_day),
         sampling_day = if_else(datetime_diff > 345600 & datetime_diff <= 432000, 
                                5, sampling_day),
         sampling_day = if_else(datetime_diff > 432000 & datetime_diff <= 518400, 
                                6, sampling_day),
         sampling_day = if_else(datetime_diff > 518400 & datetime_diff <= 604800, 
                                7, sampling_day),
         sampling_day = if_else(datetime_diff > 604800 & datetime_diff <= 691200, 
                                8, sampling_day),
         sampling_day = if_else(datetime_diff > 691200 & datetime_diff <= 777600, 
                                9, sampling_day),
         sampling_day = if_else(datetime_diff > 777600 & datetime_diff <= 864000, 
                                10, sampling_day),
         sampling_day = if_else(datetime_diff > 864000, 
                                11, sampling_day)) %>% 
  # new vars for daily means and % expected observations for each day
  group_by(area, home_winter_id, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm),
         percent_daily_obs_pm = n()/1440) %>% 
  ungroup() %>% 
  arrange(area, home_winter_id, pm_datetime_new) %>% 
  select(-first_datetime, -time_diff, -do_not_use)

@@

# write_rds(kids_pm_new4, "Output/kids_pm_new4.rds")
```

# Save cleaned dataset
```{r}
kids_pm <- read_rds("Output/kids_pm_new4.rds")
kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")

kids_pm_new5 <- kids_pm %>% 
  mutate(home_winter_id = as.character(home_winter_id)) %>% 
  select(-home, -treatment) %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id"))

# write_rds(kids_pm_new5, "Output/pm_clean.rds")
```

### 2020-01-21 home_child_data_cleaning.Rmd
data cleaning updates

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
```

>>>>

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
  mutate(winter_id = as.numeric(winter_id),
         winter_id = if_else(home == "CH218", winter_id + 1, winter_id)) %>% 
  filter(winter_id > 0)

#write_rds(kids_linked_ids, "Output/kids_linked_ids.rds")
```

### 2020-01-21 sums_data_cleaning.Rmd
data cleaning updates

>>>>

# Check to see if there are missing files to search for
```{r}
# Load SUMs data
sums_clean <- read_rds("Output/sums_clean.rds")
# Load SUMs data logs
wmt_sums_log <- read_rds("Output/wmt_sums_data_log.rds") %>% 
  mutate(area = "WMT")
nn_sums_log <- read_rds("Output/nn_sums_data_log.rds") %>% 
  mutate(area = "NN")
ak_sums_log <- read_rds("Output/ak_sums_data_log.rds") %>% 
  mutate(area = "AK")

# Combine SUMs logs from all areas
all_sums_log <- rbind(wmt_sums_log, nn_sums_log, ak_sums_log) %>% 
  mutate(winter_id = as.character(winter_id),
         sums_log = "true")

sums_check <- sums_clean %>% 
  group_by(area, home, winter_id) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  full_join(all_sums_log, by = c("area", "home", "winter_id")) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

write_csv(sums_check, "Output/sums_check.csv")
```
